---
title: "Mini-Project-CCFD"
author: "Aakarsh Sagar & Ajay Rawtani"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries required

```{r}
#Loading Relevant Libraries
library(caret)
library(rpart)
library(dplyr)
library(ROSE)
library(randomForest)
library(rpart.plot)
library(xgboost)
library(caTools)
library(ggplot2)
```

# Dataset

```{r}
# Read the CSV Data File
dat <- read.csv("creditcard.csv")
```

```{r}
#Overview of DataSet
#View the Headers
names(dat)

#View Structure of Data
str(dat)

#Print of sample set (Top 6 rows)
head(dat)

#Learn some more about the data
summary(dat)
summary(dat$Amount)
IQR(dat$Amount)
sd(dat$Amount)
var(dat$Amount)

```

```{r}
#Convert Class Column to a factor/Category {0 - no fraud; 1- fraud}
dat$Class = as.factor(dat$Class)

summary(dat$Class)
str(dat)
head(dat)
```

```{r}
# --------- PIE CHART for comparing no.of frauds and non-frauds ------------
labels = c("No Fraud","Fraud")
labels = paste(labels,round(prop.table(table(dat$Class))*100,2))
labels = paste(labels,"%")
pie(table(dat$Class),labels,col = c("#AB9968", "#BA0C2F"), main = "Overall Credit Card Transactions")
```

# Training and Testing set

```{r}
# Data split
#Check with Aakarsh on the other code (which is simpler)
set.seed(123456)

sample_data <- sample.split(dat$Class,SplitRatio = 0.80)
dat.train <- subset(dat, sample_data == TRUE)
dat.test <- subset(dat, sample_data == FALSE, select = -Class)
dat.testc <- subset(dat, sample_data == FALSE, select = Class)

dat.testc = as.data.frame(dat.testc)
colnames(dat.testc)[1] = c("Class")

# --------- PIE CHART for comparing no.of training and test ------------
train_count <- nrow(dat.train)#change to dat.train
test_count <- nrow(dat.test)#change to dat.test

train_percent <- round(train_count / (train_count + test_count) * 100, 1)
test_percent <- round(test_count / (train_count + test_count) * 100, 1)

df <- data.frame(
  dataset = c("Testing", "Training"),
  count = c(test_count, train_count),
  percent = c(test_percent, train_percent)
 )
colors <- c("#AB9968", "#BA0C2F")
labels <- paste0(df$dataset, " (", df$percent, "%)")

ggplot(df, aes(x = "", y = count, fill = dataset)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  scale_fill_manual(values = colors, labels = labels) +
  labs(title = "Count of Observations in Training and Testing Data Sets") +
  theme_void()
```

# XGBoost

```{r}
# ----------------------- XGBOOST ALGORITHM --------------------------------
labels <- dat.train$Class
y <- recode(labels, '0' = 0, "1" = 1)
xgb <- xgboost(data = data.matrix(dat.train[,-31]), 
               label = y,
               eta = 0.3,
               gamma = 0.5,
               max_depth = 6, 
               nrounds = 300, 
               objective = "binary:logistic",
               colsample_bytree = 0.8,
               verbose = 0,
               set.seed(1234)
)
xgb_pred <- predict(xgb, data.matrix(dat.test))

dat.testc$Pred = 0L
dat.testc$Pred[xgb_pred>0.5] = 1L
dat.testc$Pred = factor(dat.testc$Pred)

confusionMatrix(dat.testc$Pred,dat.testc$Class)

roc.curve(dat.testc$Class,dat.testc$Pred,plotit = TRUE,
          col="#BA0C2F",main = "ROC curve for XGBoost Algorithm",
          col.main="#AB9968")

```

```{r}
#Top contributing variables
xgb.plot.shap(data = data.matrix(dat.train[,-31]),
                                 model = xgb,
                                 top_n = 5)
```

```{r}
xgb.plot.shap.summary(data = data.matrix(dat.train[,-31]),
                                 model = xgb, top_n = 30)
```

# Logistic regression

```{r}
# ----------------------- LOGISTIC REGRESSION ------------------------------
glm_fit <- glm(Class ~ ., data = dat.train, family = 'binomial')
pred_glm <- predict(glm_fit,dat.test, type = 'response')

dat.testc$Pred = 0L
dat.testc$Pred[pred_glm>0.5] = 1L
dat.testc$Pred = factor(dat.testc$Pred)

confusionMatrix(dat.testc$Pred,dat.testc$Class)

roc.curve(dat.testc$Class,dat.testc$Pred,plotit = TRUE, col="#D6604D",main = "ROC curve for Logistic Regression Algorithm", col.main="#B2182B")
```

```{r}
summary(glm_fit)
```

# Citations

## Calculations and Formulas

1.  [Interpretation of SHAP values](https://m.mage.ai/how-to-interpret-and-explain-your-machine-learning-models-using-shap-values-471c2635b78e)

2.  [Interpretation of SHAP values alternate](https://blog.datascienceheroes.com/how-to-interpret-shap-values-in-r/)

3.  [XGBoost Mathematics Explained](https://dimleve.medium.com/xgboost-mathematics-explained-58262530904a)

4.  [XGBoost: A Scalable Tree Boosting System](https://arxiv.org/pdf/1603.02754.pdf)

5.  [XGBoost Documentation](https://xgboost.readthedocs.io/en/latest/index.html)

6.  [Interrater reliability: the kappa statistic](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3900052/#:~:text=Cohen%20suggested%20the%20Kappa%20result,1.00%20as%20almost%20perfect%20agreement)

## Dataset

7.  <https://www.kaggle.com/datasets/mlg-ulb/creditcardfraud>

-   [Andrea Dal Pozzolo, Olivier Caelen, Reid A. Johnson and Gianluca Bontempi. Calibrating Probability with Undersampling for Unbalanced Classification. In Symposium on Computational Intelligence and Data Mining (CIDM), IEEE, 2015](https://www.researchgate.net/publication/283349138_Calibrating_Probability_with_Undersampling_for_Unbalanced_Classification)

-   [Dal Pozzolo, Andrea; Caelen, Olivier; Le Borgne, Yann-Ael; Waterschoot, Serge; Bontempi, Gianluca. Learned lessons in credit card fraud detection from a practitioner perspective, Expert systems with applications,41,10,4915-4928,2014, Pergamon](https://www.researchgate.net/publication/260837261_Learned_lessons_in_credit_card_fraud_detection_from_a_practitioner_perspective)

-   [Dal Pozzolo, Andrea; Boracchi, Giacomo; Caelen, Olivier; Alippi, Cesare; Bontempi, Gianluca. Credit card fraud detection: a realistic modeling and a novel learning strategy, IEEE transactions on neural networks and learning systems,29,8,3784-3797,2018,IEEE](https://www.researchgate.net/publication/260837261_Learned_lessons_in_credit_card_fraud_detection_from_a_practitioner_perspective)

-   [Dal Pozzolo, Andrea Adaptive Machine learning for credit card fraud detection ULB MLG PhD thesis (supervised by G. Bontempi)](http://di.ulb.ac.be/map/adalpozz/pdf/Dalpozzolo2015PhD.pdf)

-   [Carcillo, Fabrizio; Dal Pozzolo, Andrea; Le Borgne, Yann-Aël; Caelen, Olivier; Mazzer, Yannis; Bontempi, Gianluca. Scarff: a scalable framework for streaming credit card fraud detection with Spark, Information fusion,41, 182-194,2018,Elsevier](https://www.researchgate.net/publication/319616537_SCARFF_a_Scalable_Framework_for_Streaming_Credit_Card_Fraud_Detection_with_Spark)

-   [Bertrand Lebichot, Yann-Aël Le Borgne, Liyun He, Frederic Oblé, Gianluca Bontempi Deep-Learning Domain Adaptation Techniques for Credit Cards Fraud Detection, INNSBDDL 2019: Recent Advances in Big Data and Deep Learning, pp 78-88, 2019](https://www.researchgate.net/publication/332180999_Deep-Learning_Domain_Adaptation_Techniques_for_Credit_Cards_Fraud_Detection)

-   [Fabrizio Carcillo, Yann-Aël Le Borgne, Olivier Caelen, Frederic Oblé, Gianluca Bontempi Combining Unsupervised and Supervised Learning in Credit Card Fraud Detection Information Sciences, 2019](https://www.researchgate.net/publication/333143698_Combining_Unsupervised_and_Supervised_Learning_in_Credit_Card_Fraud_Detection)

-   [Yann-Aël Le Borgne, Gianluca Bontempi Reproducible machine Learning for Credit Card Fraud Detection - Practical Handbook](https://www.researchgate.net/publication/351283764_Machine_Learning_for_Credit_Card_Fraud_Detection_-_Practical_Handbook)

-   [Bertrand Lebichot, Gianmarco Paldino, Wissam Siblini, Liyun He, Frederic Oblé, Gianluca Bontempi Incremental learning strategies for credit cards fraud detection, IInternational Journal of Data Science and Analytics](https://www.researchgate.net/publication/352275169_Incremental_learning_strategies_for_credit_cards_fraud_detection)
